Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-03-14 code-tests.tex
Adding section on one-zone collapse test.

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
\caption{}
\label{fig.sedov2}
\end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{1-zone free-fall test}
\label{sec.tests.freefall}
\red{(Britton)}
One zone free fall test.  Tests cooling, chemistry.

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{figures/sedov-profiles.eps}
\caption{Radially averaged profiles for the Sedov Blast Test at $t = 0.07$. Clockwise from top-left shows density, velocity, internal energy and pressure.}
\label{fig.sedov2}
\end{center}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{One-zone collapse test}
\label{sec.tests.1-zone}
The one-zone collapse test simulates the collapse of a
self-gravitating gas cloud using a semi-analytic model for the
evolution of gas density and adiabatic heat input as a function of
time.  It is designed to test the chemistry and cooling modules over a
wide range in densities and over physically motivated timescales.
Because this test disables the hydrodynamic and gravity solvers and
uses a simple model for the density evolution, it is far faster than
running a true collapse simulation.  The density evolution is based on
the self-similar Larson-Penston \citep{1969MNRAS.145..271L,
  1969MNRAS.144..425P} solution for isothermal collapse with a
modification to account for the efficiency with which the heat
introduced by compression can be radiated away
\citep{1983ApJ...265.1047Y}.  Our implementation, described briefly
here, follows the work of \citet{2005ApJ...626..627O}, which one
should consult for further details.  The density evolution is given by 
\begin{equation}
\frac{d\rho}{dt} = \frac{\rho}{t_{col}},
\end{equation}
where the collapse timescale, $t_{col}$, is
\begin{equation} \label{eqn.tcol}
t_{col} = \frac{t_{dyn}}{\sqrt{1 - f}},
\end{equation}
and $t_{dyn}$ is the dynamical time, expressed as
\begin{equation}
t_{dyn} = \sqrt{\frac{3 \pi}{32 G \rho}}.
\end{equation}
The collapse timescale is altered from the dynamical time by the
factor, $f$ in Equation \ref{eqn.tcol}, which an approximation of the
ratio of the pressure to the force of gravity.  The value of $f$
depends on the effective adiabatic index, $\gamma_{ef} \equiv (\partial
ln\ p / \partial ln\ \rho)$, which we calculate by linearly extrapolating
from derivative values for the two previous timesteps.  For the value
of $f$, we use the piecewise function of \citet{2005ApJ...626..627O},
given by
\begin{equation}
f = \left\{
  \begin{array}{ll}
  0, & \gamma_{ef} < 0.83,\\
  0.6 + 2.5 (\gamma_{ef} - 1) - 6.0 (\gamma_{ef} - 1)^{2}, & 0.83 <
  \gamma_{ef} < 1,\\
  1.0 + 0.2 (\gamma_{ef} - 4/3) - 2.9 (\gamma_{ef} - 4/3)^{2}, & \gamma_{ef} > 1.}
\end{array} \right.
\end{equation}
The specific energy evolves as
\begin{equation}
\frac{de}{dt} = -p \frac{d}{dt} \frac{1}{\rho} - \Lambda,
\end{equation}
where $\Lambda$ is the cooling rate in units of erg s$^{-1}$ g$^{-1}$
and energy, temperature, density, and pressure are related by the
ideal gas law.  Figure \ref{fig.onezone} shows an example of the
one-zone collapse test performed with an initial number density of 1
cm$^{-3}$ and temperature of 100 K using the 12 species chemistry
network with H, D, and He species and metal cooling rates calculated
with Cloudy.

\begin{figure} \label{fig.onezone}
  \begin{center}
    \includegraphics[width=1.0\textwidth]{figures/OneZoneCollapseTest.eps}
    \caption{Evolution of temperature versus number density for a one-zone
      collapse test for gas at metallicities from 0 to 10$^{-2} Z_{\odot}$
      using the primordial chemistry network with tabulated metal cooling
      rates calculated with Cloudy.}
    \label{fig.1zone}
  \end{center}
\end{figure}
